---
title: "PA1_template"
author: "DaoHengJp"
date: "2022-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r load all package}

library(knitr)
library(ggplot2)
library(lattice)
library(dplyr)
library(plyr)
library(stringi)
``` 


## 1. Code for reading in the dataset and/or processing the data

```{r activity}
activity <- read.csv("activity.csv")
activity0 <- na.omit(activity) ## dataset without NA
activity0$date <- as.Date(as.character(activity0$date, "%Y%m%d")) ## convert activity's date in DATE class
sum0 <- tapply(activity0$steps, activity0$date, sum) ## get the total steps (ARRAY class) in each day
sum0df <- data.frame(Total.steps = sum0, date = names(sum0)) ## convert sum0 into data frame by column names of sum0 and observation as sum0 total steps value
```


## 2. Histogram of the total number of steps taken each day

```{r hist1, echo=FALSE}
hist(sum0df$Total.steps, xlab = "Total number of steps", main = NULL)
```


## 3. Mean and median number of steps taken each day

```{r mean and median}

summary(sum0df$Total.steps)

```
#### The mean of steps taken each day is 10,766 / The median of steps taken each day is 10,765.


## 4. Time series plot of the average number of steps taken

```{r Time series data}

avg.step <- aggregate(steps ~ interval, activity, mean) ## aggregate mean of step in average 5mins interval

```

```{r Time series plot, echo=FALSE}
plot (avg.step, xlab = "Interval", ylab = "Number of steps", type = "l")
```

```{r max of steps}
max.avg.steps <- avg.step[ which.max(avg.step$steps),]
max.avg.steps
```
#### The interval 835 contains maximum average number of steps. 


## 5. Imputing missing values

```{r total rows of missing values}

sum(is.na(activity)) 

```
#### The total number of missing values (NA) in the dataset is 2304 rows.

```{r create new activity dataset with filling the mean of 5-minute interval into the missing value}

new.activity <- activity

for (i in 1:nrow(new.activity)) { ## range : row 1 to rows of dataset 
        if (is.na(new.activity$steps[i])) { ## condition : if the observation of step is NA , then... 
                value <- new.activity$interval[i] ## save new activity's interval value 
                steps_value <- avg.step[avg.step$interval == value, ] ## extract data frame of interval and steps by the interval value of both data set are equal  
                new.activity$steps[i] <- steps_value$steps ## assign step_value's steps to new activity 
        }
} 
```
#### New dataset with the missing value (NA) filled in 
```{r new activity dataset with the missing value filled in}

nrow(activity) == nrow(new.activity)
dim(new.activity)
head(new.activity)
```                

```{r histogram of new.activity}

new.activity$date <- as.Date(as.character(new.activity$date, "%Yd%%m"))
sum1 <- tapply(new.activity$steps, new.activity$date, sum)
sum1df <- data.frame(total.steps = sum1, date = names(sum1))
```

```{r hist2, echo=FALSE}
hist(sum1df$total.steps, xlab = "Total number of steps", main = NULL)
```

```{r mean and median of new.activity, include=FALSE}

summary1 <- summary(sum1df$total.steps)
summary2 <- summary(sum0df$Total.steps)
```
```{r comparsion of both dataset}
summary1 # Activity dataset without NA.
summary2 # Activity dataset with missing value filled in.
```
#### It is tiny diff of median bewteen two dataset. The impact of imputing missing data is able to be ignorable.

## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r weekday and weekend}

new.activity$weekday <- weekdays(new.activity$date)
val_repl0 <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
val_repl1 <- c("weekday","weekday","weekday","weekday","weekday","weekend","weekend")
new.activity$weekday <- stri_replace_all_regex(new.activity$weekday, pattern = val_repl0, replacement = val_repl1, vectorize = FALSE)
weekdaydf <- new.activity[which(new.activity$weekday == "weekday"),]
weekenddf <- new.activity[which(new.activity$weekday == "weekend"),]
avg.step.weekday <- aggregate(steps ~ interval, weekdaydf, mean)
avg.step.weekend <- aggregate(steps ~ interval, weekenddf, mean)
```

```{r plot combine}

par(mfrow=c(2,1))

```

```{r panel plot, echo=FALSE}

plot(avg.step.weekend, type = 'l', main = "weekend")
plot(avg.step.weekday, type = 'l', main = "weekday")

```                
                
                
                
        
        
        
        
        
        

